### 2017-07-02 Rotation paper/draft_2.tex
save

%\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
\shorttitle{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}
\shortauthors{Zhang\&Parsons}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
%\usepackage{rotating}
\usepackage{natbib}
%\usepackage{pdflscape}
%\usepackage{lscape}
\usepackage{ctable}
\citestyle{aa}
\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}
\graphicspath{{../figures/}}

\begin{document}

\title{Adding Sensitivity to 21cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}

%\maketitle
\author{
Yunfan Gerry Zhang \altaffilmark{1},
Adrian C. Liu \altaffilmark{1, 2},
Aaron R. Parsons\altaffilmark{1, 2}
}

\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley, CA}

\begin{abstract}
The observational effort to measure the cosmological 21cm power spectrum with radio interferometers requires high sensitivity. We introduce a straightforward and accurate method to extract sensitivity from visibility data. We develop our method in line with existing visibility-based power spectrum pipelines. Our method efficiently finds subsets of baselines to cross correlate, including partially redundant baselines, and quantifies the sensitivity contributions of given baselines. 
Using the 128-element configurations and beams of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs. We show that inclusion of non-perfectly redundant baselines 
would account for $20\%$ to $60\%$ of the sensitivity of PAPER-128 and different configurations of HERA. 
\end{abstract}

\section{Introduction}

The epoch of reionization represents the last key
stage of our universe's early evolution. Study of this event stands at
the intersection of cosmology and astrophysics. Understanding this
event not only serves as a scientific goal
of its own, but also as a gateway to crucial information
regarding fundamental physics of inflation, neutrino mass and phenomenology
of the first stars and galaxies (e.g. \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186} etc.).

Arguably the most promising observational probe of the epoch of reionization
comes from measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21cm \citep{Furlanetto2006181,PritchardLoeb}.
Other observational studies of reionization include for example Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006}. GP tests rely on Ly$\alpha$ absorption, and are thus insensitive to the degree of ionization and can only probe the end stage of reionization. The 21cm line can in comparison directly probe the
abundance and distribution of neutral hydrogen, and thus would potentially
shed light on all stages of reionization. Radio interferometric efforts
to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
Current-generation instruments include the Precision Array for Probing
the Epoch of Reionization (PAPER) \citep{Ali2015,paper32}, Murchison
Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next-generation instruments include the Hydrogen Epoch of Reionization
Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}), which is currently in planning.

One of the main challenges of observations of the cosmic
21cm line is foreground contamination. Both sources within our
galaxy, and to a lesser extent distant sources outside of our galaxy emit
radio contamination (via for example synchrotron processes) up to five orders of magnitude stronger than the
reionization signal. There are two common methods to deal with the
foreground contamination in power spectrum measurements. In the first, sometimes called  ``foreground
removal" technique, individual sources are identified and removed in the image domain.
The other technique, commonly referred to as ``foreground avoidance'', form power-spectrum directly from the visibilities, an arguably more natural approach from an instrumental point of view.  
Earlier studies have shown that the Fourier signature of smooth-spectrum foreground contaminants are constrained to a ``foreground
wedge'' \citep{wedge1,wedge2}. Contamination can thus be avoided
by restricting our observation to ``outside the wedge'', where theoretically foreground contamination is limited and noise becomes dominant. The main
challenge of using the avoidance technique is thus sensitivity. This is the motivation for the design of
the maximum redundancy arrays such as PAPER and HERA. Traditional
imaging arrays are designed to image localized bright sources and thus favor non-redundant patterns for $uv$ coverage. The redundant arrays, on the other hand,
focus on repeated measurements of the same Fourier mode
with regular grids of antennas. Since baselines
of the same length and orientation measure the same Fourier modes
on the sky, a maximum redundancy array increases the
signal to noise ratio by averaging over measurements of the same baseline.
\cite{Ali2015} provides the newest upper limit to the power spectrum
measurements with the 64-element version of PAPER (henceforth referred to as PAPER-64).

While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply simultaneously redundant baselines. Here we present a method to
improve the sensitivity output of the visibility-based power-spectrum pipelines by including pairs of partially redundant baselines. To extract maximal partial redundancy we make use of the earth's rotation. 
The earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time. In a maximally redundant array, we use the fact that baselines that are slightly different
rotate into each other at a time delay to extract power-spectrum from their respective visibilities.  Our contribution is twofold. First we introduce a formalism to include pairs of partially-redundant baselines. Secondly we show how to use the formalism to automaticaly preselect baseline pairs and time offsets, making the problem computationally efficient. Recently, \cite{wterm} proposed an algorithmic approach to extract information from partially redundant baselines, by carefully treating the $w$-term in the fringe pattern to obtain weights for cross-correlation. The algorithm makes use of a time dependent coordinate system (``moving grid'') and is thus computationally expensive. We here present a formalism that allows us to compute such weights by simply performing a numerical integral. This allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy, thereby making the power spectrum estimation from a large number of baseline pairs computationally tractable. By using pairs of baselines, we also reduce additional mode mixing. As discussed in \cite{Hazelton2013}, the varying shapes of overlapping regions in $uv$ space of all contributing baselines lead to chromaticities in the response. By considering only pairs of baselines, we fix shapes of overlapping regions, and only let their sizes vary with frequency, just as in the case of fully redundant baselines.

Baselines of the same length and orientation are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky. In order to eliminate confusion and ambiguity, we shall
introduce slightly different terminology. We shall call baselines that
are the same length and orientation ``equivalent baselines'', inspired
by the mathematical notion of equivalency classes. Two equivalent
baselines will be redundant with each other simultaneously at all
times. Non-equivalent baselines can also be partially redundant, with their partial redundancy maximized if their
respected time series is shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
another. We call these ``near-equivalent baselines''. Exactly how ``near" is ``near-enough" depends on the array configuration. All non-equivalent baselines contribute non-zero sensitivity to the power spectrum analysis It is our
goal to a) identify the near-equivalent baselines that give good
redundancy, b) to find the optimal time offset\footnote{By optimal we mean the time offset that corresponds to maximal redundancy. } for a given pair of baselines, and
c) to quantify the sensitivity improvement associated with cross multiplying
such a pair of near-equivalent baselines.

The rest of this paper is organized as follows. In section
2 we introduce the formalism for weighting partially redundant/near-equivalent baselines.
In section 3 we present numerical tests of
this technique as well as the expected sensitivity improvement
with this method for HERA and PAPER-128 and with section 4 we conclude.

\section{Method}

\subsection{Rough Idea}

\begin{figure}[H]
\includegraphics[width=\linewidth]{antpos128}

\caption{The PAPER-128 layout. Each blue dot corresponds to the location of
an antenna. Top panel shows the antenna positions drawn to scale,
bottom panel show the antenna labels and distances, excluding the outrigger
antennas.
The numbering of the antennas the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. Baselines corresponding to the same separations are called
equivalent. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines (5,69) and (5,85) are examples
of \{1, 1\} and \{1,-1\}, respectively. Note \{1, 0\} and \{-1, 0\} for example
are the same baselines and should not be counted twice.}
\label{fig:AntPos}
\end{figure}

We shall use the 128-element PAPER array to demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. We see the antenna spacing in North-South
directions are comparatively close (4m), so that baselines such as
(0, 44) and (0, 7) are very close to equivalent. In the bottom panel, the two baselines (5, 77) and (40, 85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines (5, 69) and (5, 85) are examples
of \{1, 1\} and \{1, -1\}, respectively.
Note \{1, 0\} and \{-1, 0\} for example are the same baselines and should
not be counted twice. Antennas in purely north-south baselines
are close enough to induce cross-talk, and hence are not suitable
for use. The original PAPER-64 analysis \citep{Ali2015} used three classes of baselines, here
equivalent to 
\{2, 0\}, \{2, 1\} and \{2, -1\} \cite{Ali2015} of PAPER-128. There each of these classes
of baselines are cross multiplied to itself. We shall see that in addition these
baseline classes can be cross multiplied with a time offset.

Given a point source on the sky, each baseline maps the
source to a point in $uv$ plane. As the earth rotates with respect to
the source, the point traces out tracks in the $uv$ plane. 
We show in Fig. \ref{fig:Tracks} $uv$ tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.

\begin{figure}[H]
\includegraphics[width=\linewidth]{tracks128}
\caption{Tracks of PAPER-128 (grid only, excluding outriggers) for requency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
These tracks are traced out over 0.2 sidereal days, or roughly 4.8
hours. Color represents different baselines. 
As the earth rotates, tracks are traced out counterclockwise. }
\label{fig:Tracks}
\end{figure}

Roughly speaking, we can identify
redundancy of near-equivalent baselines as crossings
of the $uv$ tracks. As we see in Fig. \ref{fig:Tracks}, there are many such crossings. Crossing tracks, however does not imply perfect redundancy. The reason lies in the finite size of the beams. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference. The inaccuracy is expected to be even more significant for the smaller HERA beam, which has a larger corresponding $uv$ point spread function.

Thus although it is a valid quick method to identify some baselines to cross-multiply, track-crossing is
not accurate enough for time offset determination, nor can give estimate of the degree of redundancy. To determine the time-offset corresponding to maximal redundancy, we must develop a more general  formalism that accounts for the point spread function of the finite beams, and estimates the degree of redundancy for general combination of baselines at a general time-offset, as we do in the next section.

\begin{figure}[H]
\includegraphics[width=1.2\linewidth]{Beams}

\caption{Sample beam response of HERA (left) and PAPER (right) antennas, both
for frequency of $\nu=150\text{MHz}$ and Stokes $I$ polarization. Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. The finite size of the beams limits the accuracy of the $uv$ track-crossing as a method of redundancy search. \label{fig:Beam}}
\end{figure}

\subsection{Formalism}
Below we shall derive theoretical expectations of cross multiplications
of two near-equivalent baselines. More precisely we shall relate
the product of visibilities of two different baselines to the power spectrum.

We take the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
dependent) beam, and $I$ is the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. $k_B$ is Boltzmann's constant. Note that $A$ here is the baseline beam, i.e. product of two antenna voltage beams, examples of which are shown in Fig. \ref{fig:Beam}. $\phi(\nu)$ is the frequency bandpass profile. In practice,
power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5). We take a moment to note that in Eq. \ref{eq:Vis1}, no flat-sky approximation has yet been made. The angular integral is performed over the dome $d\Omega$.

We define the delay-transformed visibility \citep{delay-transform}:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]
\end{aligned}
.\label{eq:Vb1}
\end{equation}
\normalsize

Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observation coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with cosmological coordinates
$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv D_{c}-Y\Delta\nu,
\end{aligned}
\]
where $\nu_{21}=1420$MHz is the 21cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}

Further we shall invoke the flat-sky limit
\begin{equation}
(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}), 
\end{equation}
where
\begin{equation}
X = \frac{c}{H_0}\int_{0}^{z}\frac{dz'}{E(z')}
\end{equation}
is a cosmological scalar. In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
the cosmological $k$ to the observation coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].

We can rewrite the delayed transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
\]
\normalsize
 where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.

We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines.

With a time offset, the beam pattern has moved relative to the sky.
Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the 3 dimensional rotation operator $\Gamma$. With implicit bounds
of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma \hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}

where in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
\begin{equation}
\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
\end{equation}
\end{widetext}
The second to third line of Eq.(\ref{eq:main}) follows from assumption of translational invariance, and the third to fourth line follows from the assumption that
the 3D power spectrum varies negligibly over the $k$-space of interest
so that $\hat{P}_{21}(k+k_{2})\approx\hat{P}_{21}(k)$. Since $\Gamma$
is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. This means that correlation is peaked at the same time-offset for all delay channels.

Finally, since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
and $\nu$, we convert the integral back to these coordinates to get
the general relation between the delay-transformed visibilities and
the power spectrum:

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
\label{eq:final}
\end{equation}

In other words the power spectrum estimate from visibilities of a baseline pair is given by 
\begin{equation}
 P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight
\begin{equation}
\Theta =\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-2\pi i\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)}. 
\label{eq:Theta}
\end{equation}

So roughly speaking the cross multiplications of visibilities at a time delay
in $uv$-space is proportional to the power spectrum times the Fourier
transform of the cross multiplied beam pattern. We can then in principle
combine information from different baseline pairs if we correct for
the phase and normalization. As a check, when applied to equivalent baselines,
$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\Gamma\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.

With Eq. \ref{eq:opp} and Eq. \ref{eq:Theta} we can for any given baseline pair and time delay, estimate the degree of redundancy, here represented by $\Theta$, thereby achieving all our goals stated in the introduction, i.e. to identify 
candidate baseline pairs with good redundancy, to find the time offset that maximizes redundancy, and to quantify the degree of such redundancy. We can do all the above simply by computing the weight $\Theta$ from
Eq.(\ref{eq:Theta}) for various time offsets, without having to actually cross-multiply visibilities at different offsets. This makes the task computationally tractable.

\subsection{Rephasing \label{sec:rephs}}
Before we discuss tests, there is one subtlety when applying Eq. \ref{eq:final} to the visibilities from correlators. The optimal time offset is given by the $\Gamma$ that maximizes the absolute value of the weight $\Theta$. However, $\Theta$ is in general complex. First integrating over the spatial dimension $\Omega$, we see the phase term at peak time-offset is inevitably frequency dependent. This frequency dependence of the phase would lead to destructive interference when we integrate over frequency.

The physical origin of this frequency-dependent phase lies in the two visibilities having different phase centers. By default the correlators phase the two visibilities  both to zenith at the same time. When they are cross-multiplied with a time lag, they must be rephased to the account for the movement of the zenith. Thus in practice we must rephase the visibilities before delay transform. Doing so is also crucial to reduction of additional mode mixing, as discussed in \cite{Hazelton2013}. Shapes of overlapping regions of pairs of baselines are fixed with respect to frequency.

In Fig. \ref{fig:freqdiff} we compare the cross-multiplied visibilities of both equivalent and near-equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz. The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair. The first and third panels show the equivalent baseline pairs \{1, 0\} against itself, and second and fourth panels show \{1, 0\} against \{1, 1\}. We see that in the un-rephased case in the second panel, although the magnitude of correlations match up for the two frequencies, the phases do not. This means summing over frequency leads to destructive interference and signal loss. The wider the frequency profile, the more destructive the interference would be. In the rephased case, the phases of the near-equivalent case match up and can be added without compromising sensitivity. We should thus rephase the data separately for each set of baseline pairs.

\begin{figure}[H]
\includegraphics[width=1.1\linewidth]{rephs}

\caption{Comparisons of the peak phases of two different frequencies. First and third panels shows equivalent baselines, second and fourth show a pair of near-equivalents. For visual simplicity only the real parts are shown. }
\label{fig:freqdiff}
\end{figure}

\section{Analysis}
\subsection{Numerical Test \label{sec:Techniquet}}

First we present some numerical tests of our formalism to verify the validity of using $\Theta$ to estimate the degree of redundancy as a function of time lag. To do so, we need to compare the amplitude and phase of the integral weight $\Theta$ 
for a pair of baselines with products of simulated visibilities of those baselines. For computational simplicity, we shall use a single frequency channel of $150$MHz for the comparison. We may use this simplification without compromising generality because Eq. (\ref{eq:final}) is valid for a range of frequency channels if and only if it's valid for each single channels, provided that different channels are combined with rephasing as described in Sec. \ref{sec:rephs}.

For the simulation, we take N=2500 random realizations of the sky
on a healpix map \citep{Heal, HealPrimer} \footnote{We use functionalities in the python package AIPY for healpix mapping
as well as coordinate transforms. }, and pick two baselines sep1,0 and sep1,1. For each realization, each pixel is given a Gaussian random value
of brightness temperature. We then rotate the baseline positions with
the appropriate rotation matrix, multiplying the sky by the primary beam to get the visibilities, for each
baseline\footnote{There are two obvious ways to achieve the rotation. One
can either fix the sky and rotate the baselines, or the other way
around. We found however, that we must not physically rotate the sky
map, for the numerical round-offs due to finite resolutions of the
map turns out to be significant. Thus we let the sky, represented
by the healpix map, be fixed, and rotate the baselines. }. The resulting visibilities for the two baselines are then convolved
via the Fourier convolution theorem, to obtain values of the cross
correlation as a function of time-offset. The peak of the curve then
corresponds to maximum redundancy.  We do this for both the equivalent (sep1,0 against sep1,0) and near-equivalent (sep1,0 against sep1,1) case. The accuracy of this result is
limited by (simulated) cosmic variance and finite spacial resolution,
and hence can be beat down by averaging over a large number of universes.
A numerical estimate of the error of the peak height with $N=1$ is
$\lesssim20\%$, and thus with $N=2500,$ we achieve an error of peak
height $<0.5\%$.

\begin{figure}[h]
\includegraphics[width=\linewidth]{opp}

\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}). We generated 2400 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.  Top panel shows the equivalent baseline pairs sep1,0 against sep1,0, bottom panel shows set1,0 against sep1,1.  Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines. In both cases blue denotes real part, green the imaginary part, and red the magnitude. We see that in both cases the theory and simulation line up in both amplitude and phase.}
\label{fig:numerics}
\end{figure}

The comparison is show in Fig. \ref{fig:numerics}. 
Sensitivity contribution is inferred from height of the peak of the cross-multiplied visibilities as a function of time. 
We first compare the simulation to the analytical weight $\Theta$ for a pair of equivalent baselines (sep1,0, sep1,0), then normalize the peaks 
of the near-equivalent baselines (sep1,0, sep1,1) to that of the equivalent ones. 
In Fig. \ref{fig:numerics}  we show on the top comparison of
the convolution of an equivalent baseline of PAPER-128, normalized to 1 at the
peak, whose location is at zero time offset as expected. On the bottom we show
comparison of a near-equivalent baseline pair, of classes sep1,0:1,1 and sep1,1:1,1. The peaks
are normalized with the same factor as the peak height on the top,
i.e. the two plots have thus the same scale on the vertical axis.
The height of the plot thus quantifies the added sensitivity. We see at a peak of around 0.055 sidereal days, or 1.32 hours, the two baselines are maximally redundant.

At the optimal time separation, the integral in Eq. \eqref{eq:final}
is maximized. Thus as another check we expect the two beams to have to same fringe pattern
(frequency and phase). Due to the time delay, however, the beam center
would be slightly shifted with respect to each other. This we show
in Fig. \ref{fig:Beam-fringe-pattern}. The left and middle panels show the beam fringe
patterns for baselines sep2,0 and sep2,1, delayed by 0.0325 sidereal days,
and the right panel shows their cross product. The fringe pattern
indeed cancels out as we expect.

\begin{figure*}[h!]
\includegraphics[width=0.9\textwidth]{fringe_res}
\caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
and their conjugate product (right). Frequency of $\nu=0.15\text{GHz}$ is
chosen and only the real components are shown. The colorbar values are normalized such that the peak of the original beam is 1. \label{fig:Beam-fringe-pattern}}
\end{figure*}

\begin{figure*}[h!]
\includegraphics[width=0.9\textwidth]{sensitivity}

\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128. In the legend m,n:p,q denotes cross
multiplying PAPER-128 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
panel shows the peak height (degree of correlation) of each baseline
combination, while the bottom panel multiplies the heights by the
corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of $\sqrt{2}$ (explained in text). 
In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
contribution of 1,0:1,0 to unity. The same symbols and colors correspond in the top and bottom panel, with half of the legend shown in each. }
\label{fig:sensplot}
\end{figure*}

\begin{figure*}[h!]
\includegraphics[width=0.9\textwidth]{HeraAntpos}
\label{fig:HeraAntpos}
\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations. HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases. For HERA350, only the 320 elements in the core are shown. }
\end{figure*}

\subsection{ Sensitivity \label{sec:sensitivity}}

Having verified Eq. \eqref{eq:final}, we can thus predict the sensitivity contributions of a
particular baseline pair simply by computing the integral Eq. \.
Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam. A larger beam like that of PAPER would tolerate larger time-offsets because more sky area can coincide in the two beams \footnote{Larger beams also imply smaller spread in $uv$ space. This could lead to either larger or smaller redundancy, which depends on the overlap of two such point spread functions.}
Having computed all of the baseline pairs, we find that baseline pairs that are mirror images of each other 
give the same amount of redundancies (peak height), with the opposite time offset, as expected from symmetry. 
For example, sep1,0:1,1 is mirror image of sep1,0:1,-1 and these two baseline pairs
give the same sensibility contribution. Thus we shall only show a subset of representative baseline pairs to illustrate the contributions from different classes of baseline pairs. For a more complete result see Fig. \ref{fig:pairplot}.

In the top panel of Fig. \ref{fig:sensplot}
we show the peak heights and locations for a variety of baseline combinations.
We see that baseline pairs that have crossings at a smaller time delay
tend to have higher correlations. In other words, correlation peaks
that are closer to zero time lag are higher. This is expected since
a) the longer the time delay, the more the antennas have moved with respect
to the sky and hence the less overlaps in patch of sky surveyed, b) smaller optimal
time-offset corresponds to smaller differences in orientation, and hence in length of
baselines in PAPER.

To determine that actual relative contribution to sensitivity of these
baseline pairs, we have to take into account of the multiplicities of
these baselines. By these we mean how many physical antenna pairs have the
same length and orientation. Looking at Fig. \ref{fig:sensplot}
we see for example sep1,0 will have higher multiplicity than sep2,0,
or sep1,1. The latest release of PAPER-64 data uses the 128-equivalent baselines sep2,1,
sep2,0 and sep2,-1 \citep{Ali2015}, and achieved a $2\sigma$ upper
limit of $(22.4\text{mK})^{2}$. There, the three sets of equivalent baselines
are only cross multiplied by itself. Assuming that each baseline delivers
the same quality of data (meaning they have the same height of correlation
peaks, which is in our normalization equal to unity), the relative
contribution to sensitivity can be estimated.

First we can average of the visibilities of the equivalent baselines. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
are $M=(16-|m|)\times(7-|n|)$ copies of the baseline sepm,n. This means
that if we add visibility measurements of all these equivalent baselines,
we get a factor of $\sqrt{M}$ reduction in noise
level $\sigma_N$ of the visibility. The sensitivity contribution of sepm,n, cross multiplied with sepm',n'  thus roughly speaking scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
For cross-multiplications of near-equivalent baselines of
types sepm,n and sepm',n', we get an effective weight:

\begin{equation}
\widetilde{\Theta}_{bb'} \propto \Theta_{bb'}\times\sqrt{MM'}.\label{eq:sensul}
\end{equation}

Shown in the bottom panel of Fig. \ref{fig:sensplot} is the peak heights weighted
by the multiplicity factor. Points that have zero time delay are the equivalent baseline pairs and their weighted correlation values simply reflect the multiplicity factor. For clarity of presentation we have ``folded over'' the negative time delays and combined baseline pairs that are identical modulus parity, 
or in other words are mirror images of each other
%\footnote{The parity symmetry is partially broken unless the array is located at the poles. But we are ignoring this here for illustrative purposes. See Fig. \ref{fig:pairplot} for a more accurate result. }
, by summing over them to get an extra factor of $\sqrt{2}$. For example, 
instead of plotting 1,1:1,1 and 1,-1:1,-1 separately, we plot 1,1:1,1 with twice the multiplicity. Similarly, baseline pairs such as 
1,0:1,1 also get the factor of 2 because they are identical to 1,0:1,-1. Baseline pairs such as 1,0:1,0, or 1,1:1,-1 do not get
the factor of $\sqrt{2}$ because their mirror images are themselves.

Having defined the modified weight $\widetilde{\Theta}$, we can estimate the power spectrum by inverse covariance weighting:
\begin{equation}
\begin{aligned}
 P(k_{\tau}) &= \frac{\sum_{bb'}P(k_{b,\tau})/\sigma_P^2(bb')}{\sum_{bb'}\sigma_P^2(bb')}, \\
 &= \frac{\sum_{bb'}P(k_{b,\tau})\widetilde{\Theta}_{bb'}^2}{\sum_{bb'}\widetilde{\Theta}_{bb'}^2},
 \end{aligned}
\end{equation}
where the sum is over classes of baseline pairs. 
We define the estimator sensitivity to be the inverse of the power-spectrum noise variance:
\begin{equation}
\rho \propto \frac{1}{\sigma_P^2} \propto \rho_0^2\sum_{bb'}^N\widetilde{\Theta}^2_{bb'},
\end{equation}
where, if $\sigma_S^2$ and $\sigma_N^2$ are the characteristic signal and noise levels of a single-baseline visibility, $\rho_0=\sigma_S^2/\sigma_N^2$ is the signal to noise ratio.

The scaling in Eq. \ref{eq:sensul} was rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional $\rho_0$:

\begin{equation}
\label{eq:tildereal}
\widetilde{\Theta}_{bb'}=\frac{\Theta_{bb'}\sqrt{M_bM_{b'}}}{\sqrt{1 + \rho_0 \left(M_b+M_{b'} \right)}}.
\end{equation}

For a given $\rho_0$ Eq.\ref{eq:tildereal} thus quantifies the relative sensitivity contribution of a baseline pair $bb'$. Assuming a reionization signal of $\Delta_{21cm}^2\sim 30mK^2$, observation at 150MHz ($z=8.5$), 120 days of observation with PAPER antennas, we have roughly
(See Eq.(20) in \cite{first-paper})
\begin{equation}
\rho_0 \sim 0.001\left[\frac{L}{40m}\right] \left[\frac{0.1hMpc^{-1}}{k}\right]^3, 
\end{equation}
where $L$ is the baseline length. We only need a single characteristic $L$ even in the near-equivalent case because only baselines of nearly equal length would have high redundancy.As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.

\subsection{Array Configuration Comparisons \label{sec:arrconf}}
We run our algorithm over all possible baseline-pairs of  PAPER128, HERA37, HERA128, HERA243 and HERA350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos}. The  
hexagonal design is the densest pattern of antenna-packing. The larger arrays are designed with a ``gap'' dividing the antennas into three different regions. The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more near-equivalent baselines than the versions without the gap. The motivations behind the designs are explained in \cite{HERAconfiguration}.  The lack of short baselines that are visually close to each other, as well as the smaller beam (Fig. \ref{fig:Beam}) means that we expect to see only longer near-equivalent baselines. The lower multiplicities per class of baselines is made up with the larger number of classes of baseline-pairs, especially given the gap in the larger versions.

Having quantified the sensitivity from a given pair of baselines, we study the cumulative sensitivity of the array depending on which baseline pairs we include. Evidently we should prefer the pairs with larger $\widetilde{\Theta}$. In Fig. \ref{fig:osens} we plot $\rho$ against the minimum $\widetilde{\Theta}$. $\rho(\widetilde{\Theta}_{min})$ is the sensitivity of the array when baseline-pairs that have $\widetilde{\Theta}>\widetilde{\Theta}_{min}$ are included. The dashed lines represent the values when only the equivalent baseline-pairs are used.  We see as expected that in all cases using the near-equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when worse baseline pairs are used. The small HERA37, with no gap (like in HERA350) or short near-equivalent baselines (like in PAPER 128), will not benefit much from the near-equivalent baselines. The maximum benefits for other cases are expected to be around $20\%$ to $60\%$. PAPER128 is designed with highly redundant near-equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from near-equivalent baselines at low $\widetilde{\Theta}_{min}$ due to there being more classes of such pairs. Note that here we normalized $\rho$ such that the contribution of the top equivalent baseline pair, such as (sep0,1:0,1 in the PAPER128 case) are 1. This plot therefore does not compare the absolute sensitivity across the different arrays. The stepwise pattern is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in ``batch''. 
%\begin{centered}
\begin{figure}[H]
\includegraphics[width=\linewidth]{osens}
\caption{Sensitivity of redundant arrays as a function of the minimum effective weight. Dashed lines represent when only equivalent baseline pairs are used, while solid lines indicate use of both the equivalent and near-equivalent baselines are used. The y-axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison of absolute sensitivity across the different arrays shown.  }
\label{fig:osens}
\end{figure}
%\end{centered}

%\subsection{Notes on Data, Foreground and Noise}
%Unlike the simulated clean EOR signal, real data are dominated by foreground %and thermal noise, neither of which turns out to exhibit the correlation %pattern seen in Fig.  \ref{fig:numerics}. Foreground exhibits long-range %correlations and would lead to high responses at wide range of time-offsets. %Instrumental noise for different baselines/antenna-pairs, on the other hand, are uncorrelated and thus exhibit responses consistent with zero at all time offsets. 
%To illustrate, we use the second observing season of PAPER-128 data, taken from Jan. 21st to March 7th of 2014. To illustrate the effect of noise and foreground, we use the data before fringe-rate and delay filtering.

\section{Conclusion}
Redundant arrays are designed to maximize sensitivity. Current generations of redundant radio arrays, such as those probing the power spectrum of the epoch of reionization could benefit from data analysis techniques that improve the sensitivity. We present an intuitive analysis of cross-multiplying baselines that are close in length and orientation to each other. Given an antenna array configuration, our method quickly identifies the best baseline pairs to cross-multiply and predict the expected sensitivity contribution. With the predicted result one can improve existing power-spectrum pipelines through 3 simple steps. 1). Rephase the visibilities prior to delay transforming by the zenith displacement (or more accurately by the phase predicted by our numerical analysis). 2). Shift the visibilities in time. 3) Cross multiply the visibilities of the two baselines  to form the power-spectrum. 4) Finally combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case. We showed that such techniques could lead to $20\%$ to $60\%$ increase in sensitivity for PAPER and HERA.

\pagebreak

\appendix
\section{\label{sec:appB}\\Derivation of Noise Covariance \label{sec:appB}}
\label{sec:appB}
In this appendix we give a brief derivation of the effective weight quoted in \ref{sec:sensitivity}. We combine the different power spectrum measurements by inverse variance weighting \footnote{In practice techniques such as bootstrapping is often used, see for example \citep{Ali2015}}. We shall separate the visibility and power spectrum into signal and noise contributions:
\begin{equation}
\begin{aligned}
V &= V_S+V_N,\\
P &= P_S+P_N.
\end{aligned}
\end{equation}
We shall denote the noise variance of power spectrum and visibility
\begin{equation}
\begin{aligned}
\sigma_V^2 &= \langle |V_N|^2 \rangle,\\
\sigma_P^2 &= \langle P_N^2 \rangle.
\end{aligned}
\end{equation}
One may notice that we have used a single covariance for the complex quantity visibility. It's simple to show that the same result holds if we use a separate real and imaginary components, as long as they are independent of each other. In fact, for simplicity and without loss of generality we shall treat the visibility as a real quantity in the rest of this derivation. 
Note that though we can assume $\langle V_N^{odd-power}\rangle=0$, the same is not true for $P_N$.

Then the variance of $P$ constructed with visibilities $V_1$ and $V_2$ from two baseline classes can be estimated\footnote{We assume all noise terms to be independent for simplicity, in practice the correlation of different measurements ifrom equivalent baselines are alleviated by grouping the baselines in the class and the days of observation, as in \cite{Ali2015}}:
\begin{equation}
\begin{aligned}
\sigma_P^2 &= \langle P^2\rangle -\langle P \rangle^2,\\
&\propto \langle \frac{(V_{1S}+V_{1N})^2 (V_{2S}+V_{2N})^2}{\Theta^2} \rangle - \langle \frac{(V_{1S}+V_{1N}) (V_{2S}+V_{2N})}{\Theta} \rangle ^2,\\
&= \frac{1}{\Theta^2} \left( V_{1S}^2\sigma_{V2}^2+V_{2S}^2\sigma_{V1}^2+\langle V_{1N}^2 V_{2N}^2\rangle\right), \\
&= \frac{1}{\Theta^2} \left[ V_{S}^2(\sigma_{V2}^2+\sigma_{V1}^2) + \sigma_{V1}^2 \sigma_{V2}^2\right], 
\end{aligned}
\end{equation}

where in the second last line we have substituted visibility noise variance. In the final line we used Wick's theorem and the fact that the signal from two visibilities are equal.

Recall from the discussion on multiplicities we can write
\begin{equation}
\sigma_V^2=\frac{\sigma_0^2}{M},
\end{equation}
where $\sigma_0$ is some single-baseline noise level. Letting $\rho_0=V_S^2/\sigma_0^2$ be the signal to noise ratio for a single baseline, we can write

\begin{equation}
\begin{aligned}
\sigma_P^2 & \propto  \frac{\sigma_0^4}{\Theta^2} \left[ \rho_0 \left(\frac{1}{M_1}+\frac{1}{M_2} \right) + \frac{1}{M_1 M_2}\right], \\
&\propto \frac{1}{\widetilde{\Theta}_{12}^2},
\end{aligned}
\end{equation}
where we have defined a slightly modified version of the effective weight (compare with Eq. \ref{eq:sensul}):
\begin{equation}
\widetilde{\Theta}_{12}=\frac{\Theta_{12}\sqrt{M_1M_{2}}}{\sqrt{1 + \rho_0 \left(M_1+M_{2} \right)}}.
\end{equation}

\section{\label{sec:fullstats}\\Additional Statistics}

\begin{widetext}
\begin{figure}[H]
\includegraphics[width=\textwidth]{pairplot}

\caption{Pairplots of the top contributing baseline pairs in three arrays. Plotted properties are optimal time delay $dT$, peak height $\Theta$, effective weight $\widetilde{\Theta}$ and baseline length $L$. Only those points with $\widetilde{\Theta}>0.05$ are shown (the weight for the top class of equivalent pair is normalized to 1). Only one baseline length is shown since all top contributing pairs have very similar lengths as expected. The scatter plots are shown with transparency so that darker regions indicate degeneracies. Scatter points of the 3 arrays overlap, in order indicated by the legend. }
\label{fig:pairplot}
\end{figure}
\end{widetext}

In Fig. \ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.05$, where again the weight for the top class of equivalent pair is normalized to 1). The $dT$ vs. $\Theta$ and $dT$ vs. $\widetilde{\Theta}$ plots are familiar from Fig. \ref{fig:sensplot}. We only show 3 of the mentioned arrays for visual clarity. The other two results are similar barring intuitive differences. We make a few observations:
\begin{itemize}

\item Mapping of equivalent baseline pairs: the equivalent baselines are located in key points in the plots. They are the vertical bars in the lower left four plots and the highest peak in the first three histograms, the upper "diagonal edge" in the $\widetilde{\Theta}$ vs. $dT$ and the dot of $\Theta=1$.

\item Longer baselines roughly peak at lower time-offset, and lower multiplicity. This is intuitive for PAPER since  for the same north-south separation the rotation angle is smaller with longer east-west separation.

\item Interestingly for HERA the near-equivalent baselines form multiple "U" shapes in $dT, \Theta$ plane, somewhat counter-intuitive especially given out last point. This is characteristic of the hexagonal layout. Baselines of equal length are more redundant at the right time delay than ones that are slightly different in length, though the latter has a shorter optimal time-delay.

\item The top near-equivalent pairs in HERA the same $\Theta$ as in PAPER128, but much lower $\widetilde{\Theta}$. This is because they are longer baselines with lower multiplicity. In the end these baseline classes still lead to high contributions to total sensitivity (Fig. \ref{fig:osens}) because there are a lot more such baseline pairs for HERA.

\end{itemize}

\bibliographystyle{apj}
%\nocite{*}
\bibliography{draft_working}

\end{document}